---
title: "Lab 1 - Intro to R and RStudio - KEY"
author: "Total: 90 pts"
date: ""
output: 
  html_document: 
    highlight: pygments
    theme: cerulean
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

* * *

## Lab report

**Load Arbuthnot data**
```{r}
load(url("https://stat.duke.edu/~mc301/data/arbuthnot.RData"))
```

#### Exercise 1:

~~~
5 pts - no narrative needed
~~~

Either of the following is considered correct:

```{r}
arbuthnot$girls
```

```{r}
arbuthnot %>%
  select(girls)
```

#### Exercise 2:

~~~
15 pts:
* 7 pts for plot
* 8 pts for interpretation
~~~

* The overall trend of number of girls baptized over years is upwards. (2 pts)
* However there is a sharp dip between 1640 - 1650. Baptisms of girls pick up again in 1660. (4 pts)
* The rate of increase slows down around 1690. (2 pts)

```{r}
qplot(x = year, y = girls, data = arbuthnot, geom = "line")
```

#### Exercise 3:

~~~
20 pts:
* 6 pts for creating the new variables
* 6 pts for plot
* 8 pts for interpretation
~~~

* The proportions of boys baptized each year are slightly over 50%. (4 pts)
* There are cyclical patterns but overall there seems to be a slight decrease in the proportions over time. (4 pts)

```{r fig.width=7, fig.height=5}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls) %>%
  mutate(boy_ratio = boys / total)

qplot(x = year, y = boy_ratio, data = arbuthnot, geom = "line")
```

* * *

### More practice

**Load present day data**
```{r}
load(url("https://stat.duke.edu/~mc301/data/present.RData"))
```

#### 4:

~~~
15 pts
~~~

* The years included in the dataset are 1940 - 2013. (5 pts)
* The dataframe has 74 observations and 3 variables. (5 pts)
* The variables are `year`, `boys` (number of boys born), and `girls` (number of girls born). (5 pts)

#### 5:

~~~
5 pts
~~~

The counts are much higher than in the Arbuthnot data set, they are not on the same scale.

#### 6:

~~~
20 pts:
* 5 pts for creating the new variables
* 5 pts for plot
* 10 pts for interpretation
~~~

* Arbuthnot's observation about boys being born in greater proportion than girls holds up in the U.S. as well. (5 pts)
* However while the ratio of boys is much higher than girls in the 1940s, the ratio decreases and approaches 50% throughout time. (5 pts)

```{r fig.width=7, fig.height=5}
present <- present %>%
  mutate(total = boys + girls) %>%
  mutate(boy_ratio = boys / total)

qplot(x = year, y = boy_ratio, data = present, geom = "line")
```

#### 7: 

~~~
10 pts:
* 8 pts for code
* 2 pts for answer in text
~~~

We saw the most number of births in 2007.

```{r}
present %>%
  arrange(desc(total))
```
