---
title: "Lab 3 - Probability - KEY"
author: "Total: 90 pts"
date: ""
output: 
  html_document: 
    highlight: pygments
    theme: cerulean
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```


**Load Hot Hand files:**
```{r}
load(url("https://stat.duke.edu/~mc301/data/hot_hand.RData")) 
```

## Exercises: (50 pts)

#### Exercise 1: 

~~~
10 pts - 5 pt for each

~~~

A streak length of 1 means a hit followed by a miss. 

A streak length of 0 is just a miss (which must occur after a miss that ended the preceeding streak).

#### Exercise 2:

~~~
15 pts 
* 5 pts for `kobe_streak`
* 5 pts for plot
* 5 pts for intepretation
~~~

The distribution of Kobe's streaks is unimodal and right skewed. The typical length 
of a streak is 0 since the median of the distribution is at 0. The IQR of the 
distribution is 1. The longest streak of baskets is of length 4. Streak lengths of 
3 and 4 are unusually high compared to the rest of the distribution.

```{r plot-streak-kobe}
kobe_streak <- calc_streak(kobe_basket$shot)
qplot(data = kobe_streak, x = length, geom = "histogram", binwidth = 1)
```

#### Exercise 3: 

~~~
15 pts 
* 5 pt for `sim_unfair_coin`
* 10 pt for `table(sim_unfair_coin)`)
~~~

81 tails and 19 heads (Note: Results will vary. A seed of `9102015` indicates they probably did not follow the directions.)

```{r set-seed}
set.seed(9102015)
coin_outcomes <- c("heads", "tails")
sim_unfair_coin <- sample(coin_outcomes, size = 100, replace = TRUE, 
                          prob = c(0.2, 0.8))
table(sim_unfair_coin)
```

#### Exercise 4: 

~~~
10 pts 
* 4pt for correct size
* 3pt for correct probablity
* 3pt for assigning to correct object `simbasket`
~~~

```{r sim-basket-prob}
shot_outcomes <- c("H", "M")
sim_basket <- sample(shot_outcomes, size = 133, replace = TRUE,
                     prob = c(0.45, 0.55))
```

* * *

## More practice (40pts):

#### 1
~~~
8pts
~~~

```{r}
sim_streak <- calc_streak(sim_basket)
```

#### 2
~~~
8 pts 
* 4 pts for intepretation
* 4 pts for plot
~~~

The distribution of simulated streaks is also unimodal and right skewed. The 
typical length of a streak is 0 since the median of the distribution is at 0. The 
IQR of the distribution is 1.25. Streak lengths of 4 and above are unusually high 
compared to the rest of the distribution. (Note: Answers will vary.)

```{r}
qplot(data = sim_streak, x = length, geom = "histogram", binwidth = 1)
sim_streak %>%
  summarise(median(length), IQR(length))
```

#### 3: 
~~~
8 pts - Should be similiar, but not exactly the same.
~~~ 

We would expect the distribution to be somewhat similar, i.e. unimodal and right 
skewed with a median around 0. But we wouldn't get the exact same streaks, and therefore the distribution will not be exactly the same.

#### 4: 
~~~
11 pts 
* 5 pt for similar
* 6 pt for no evidence for hot hand
~~~

The distributions look very similar. Therefore, there doesn't appear to be 
evidence for Kobe Bryant's hot hand.

#### 5: 
~~~
5 pts - something reasonable
~~~

In lecture and homework problems we have seen the concepts of independence and 
dependence as well as conditional probabilities. Simulations are something we've 
seen in a class activity in lecture. (Note: Answers will vary.)
