---
title: "Introduction to data"
output:
  html_document:
    css: ../lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
library(dplyr)
library(ggplot2)
```

Some define statistics as the field that focuses on turning information into knowledge. The first step in that process is to summarize and describe the raw information - the data. In this lab we explore flights, specifically a random sample of domestic flights that departed from the three major New York City airport in 2013. We will generate simple graphical and numerical summaries of data on these flights and explore delay times. As this is a large data set, along the way you’ll also learn the indispensable skills of data processing and subsetting.

## Getting started
### Load the packages
Same as last time, we will use the `dplyr` and `ggplot2` packages.
```{r}
library(dplyr)
library(ggplot2)
```

### Download the lab report template
We created a lab report template for you this time as well. However, you need to put in the code needed for each exercise by yourself (by copying from the `intro_to_data.html` file and perhaps adding your own code).


