---
title: "Lab 4 - Sampling distributions - KEY"
author: "Total: 90 pts"
output: 
  html_document: 
    highlight: pygments
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
library(dplyr)
library(ggplot2)
library(mosaic)
```

* * *

#### Load data:
```{r load-data}
load(url("https://stat.duke.edu/~mc301/data/ames.RData"))
```


#### Set a seed: 
~~~
2 pts for setting some seed
~~~

```{r set-seed}
set.seed(2102016)
```

## Exercises:

#### Exercise 1: 
~~~
10 pts

- 2 pt for plot
- 2 pt for summary stats
- 6 pts for description: 
  - 2 pt for using median & IQR instead of mean & sd
  - 2 pt for mentioning Q1 and Q3 as suggested in the instructions
  - 2 pt for stating min and max 
~~~

The distribution of areas of houses in Ames is unimodal and right-skewed, with a 
median of 1,442 square feet. The middle 50% of the houses range between 1,126 
square feet and 1,743 square feet, yielding an IQR of 617 square feet. The 
smallest house is 334 square feet and the largest is 5,642 square feet.

```{r fig.width=7, fig.height=5}
qplot(data = ames, x = area, binwidth = 250)
ames %>%
  summarise(mu = mean(area), pop_med = median(area), 
            sigma = sd(area), pop_iqr = IQR(area),
            pop_min = min(area), pop_max = max(area),
            pop_q1 = quantile(area, 0.25),
            pop_q3 = quantile(area, 0.75)) 
```


#### Exercise 2: 

~~~
18 points 

Note: Answers may vary since randomly sampling.

- 2 pt for sampling code
- 2 pts for plot
- 6 pts for summary stats: 3 pt for basic code, 3 pt for not using parameter as 
labels (like x_bar instead of mu)
- 6 pts for description: 2 pt for using median & IQR instead of mean & sd, 2 pt for mentioning Q1 and Q3 as suggested in the instructions, 2 pt for stating min and max 
- 2 pt for reasonable comparison to population distribution
~~~

The distribution of areas of houses in this sample is unimodal and right-skewed, 
with a median of 1,404 square feet. The middle 50% of the houses in this sample range 
between 1,014 square feet and 1,685 square feet, yielding an IQR of 672 square feet. 
The smallest house in the sample is 630 square feet and the largest is 3,140 square feet.

The population and sample distribution are similar in shape but the sample d
distribution appears to be slightly less skewed.

```{r fig.width=7, fig.height=5}
samp1 <- ames %>%
  sample_n(50)
qplot(data = samp1, x = area, binwidth = 250)
samp1 %>%
  summarise(xbar = mean(area), samp_med = median(area), 
            s = sd(area), samp_iqr = IQR(area),
            samp_min = min(area), samp_max = max(area),
            samp_q1 = quantile(area, 0.25),
            samp_q3 = quantile(area, 0.75))
```


#### Exercise 3: 
~~~
4 pts

- 2 pt: not same
- 2 pt: somewhat different
~~~

We would not expect the means of the two samples to be exactly the same (unless the
same seed was used for sampling), but we would expect them to just be somewhat 
different and not very different.

#### Exercise 4: 

~~~
6 pts

Note: First part of the answers may vary since randomly sampling.

- 2 pt for code/output for sampling and calculating mean of `samp2`
- 2 pt for comparing two samples with n = 50
- 2 pt for choosing between n = 1000 and n = 100 
~~~

The means of the two samples with 50 observations are close to each other.

We would expect a sample with 1000 observations to yield a sample mean that would be a more accurate estimate of the population mean than a sample with 100 observations.

```{r}
samp2 <- ames %>%
  sample_n(50)

samp2%>%
  summarise(x_bar = mean(area))
```


#### Exercise 5: 

~~~
6 pts 

- 2 pt for code / output
- 2 pt for 2,000 elements
- 2 pt for distribution description
~~~

There are 2,000 elements in `sample_means50`. The distribution of this sample is 
unimodal and symmetric, with mean 1,499 square feet and standard deviation 68.88 
square feet. 

```{r}
sample_means50 <- do(2000) * ames %>% 
                              sample_n(50, replace = TRUE) %>%
                              summarise(x_bar = mean(area))

qplot(data = sample_means50, x = x_bar)
```


#### Exercise 6: 

~~~
6 pts

- 2 pt for code / output
- 2 pt for 25 observations
- 2 pt for each observation is a sample mean
~~~

The object `sample_means_small` contains 25 elements, and each observation 
represents a mean from a simple random sample.

```{r}
sample_means_small <- do(25) * ames %>%
                                sample_n(10, replace = TRUE) %>%
                                summarise(x_bar = mean(area))
sample_means_small
```

#### Exercise 7: 

~~~
10 pts

- 2 pts for observations are sample means
- 2 pts for shape
- 2 pts for center
- 2 pts for spread
- 2 pts for nothing changes when number of simulations is increased
~~~

Each observation in the sampling distribution represents a sample mean.

As the sample size increases, the shape of the distribution becomes more
symmetric (approaches normal), the standard error decreases, but the mean 
stays roughly the same.

Increasing the number of simulations does not affect the distribution.

* * *

## More Practice:

#### 8. 

~~~
4 pts 

Note: Answers may vary since randomly sampling.

- 2 pt for code
- 2 pt for estimation
~~~

The best point eastimate is the mean of the sample, which is $216,608.

```{r}
samp1 <- ames %>%
  sample_n(15)

samp1%>%
  summarise(x_bar = mean(price))
``` 


#### 9 

~~~
10 pts 

Note: First part of the answers may vary due to random sampling.

- 3 pt for sampling code
- 2 pt for plot code/output
- 2 pt for estimation
- 2 pt for confming with population mean
~~~

The distribution is unimodal and slightly right skewed, with mean $180,129 and 
standard deviation $20,608. Based on this sampling distribution the best guess for 
the population mean is $180,129, which is pretty close to the population mean of 
$180,796.

```{r fig.width=7, fig.height=5}
sample_means15 <- do(2000) * ames %>%
                              sample_n(15, replace = TRUE) %>%
                              summarise(x_bar = mean(price))
                          
qplot(data = sample_means15 , x = x_bar)

sample_means15 %>%
  summarise(mean = mean(x_bar), sd = sd(x_bar))

ames %>%
  summarise(pop_mean = mean(price))
```


#### 10 

~~~
10 pts

Note: First part of the answers may vary due to random sampling.

- 3 pt for sampling code
- 2 pt for plot code/output
- 2 pt for estimation
- 2 pt for comparisons
~~~

The distribution is unimodal and symmetric with mean $180,821 and a standard deviation
of 6,580. The mean of this sampling distribution is closer to the population mean 
than the mean of the previous sampling distribution, the shape is more symmetric than 
the previous distribution. Based on this distribution we would update our best guess 
for the population mean to be $180,821.

```{r}
sample_means150 <- do(2000) * ames %>%
                                sample_n(150, replace = TRUE) %>%
                                summarise(x_bar = mean(price))
                          
qplot(data = sample_means150, x = x_bar)

sample_means150 %>%
  summarise(mean = mean(x_bar), sd = sd(x_bar))
```


#### 11 

~~~
4 pts

- 2 pt for which has smaller spread
- 2 pt for which to prefer
~~~

The standard error (measured as the standard deviation of the sampling distribution)
for the second distribution is smaller than the first, which makes sense since the 
sample size is higher for the second sample. If we're concerned with making estimates 
that are more often close to the true value, would we prefer a distribution with a 
small spread.
